Holomorphic field theories and higher algebra

### 2024-10-01 draft.tex
just trucking along

Before jumping into the main story, we mention a motivation for talking about $E_d$- and $\PDisc_d$-algebras in parallel.
By dealing with both, we are suggesting, of course, that there is a natural analogy and hence a useful transfer of intuition between the settings of topological and holomorphic field theories.
But there is a stronger, technical relationship that we'll develop via a version of dimensional reduction.
From each holomorphic field theory in complex dimension $d$, 
we'll extract an $E_d$-algebra.
This construction generalizes the relationship between the Kac-Moody vertex algebra of a Lie algebra $\fg$ at level $k$ and the central extension of the loop algebra~$L\fg$.
We expect this correspondence to bear rich fruit.
\owen{that metaphor's a bit fruity}

There is an explicit model for an algebra over the cohomology of the $E_d$-operad called a $P_d$-algebra.
Essentially, these are graded commutative algebras with a Poisson bracket of cohomological degree $-d+1$.
In \cite{CostelloScheimbauer} a parallel model for the cohomology of the holomorphic $\PDisc_d$ colored operad was proposed; though the exact axioms were not written down.
In recent work, similar models to the one given in \cite{CostelloScheimbauer} have appeared using so-called (multivariable) `$\lambda$-brackets' \cite{BGKWY,GKW}. This work uses such an algebraic structure to give a beautiful proposal for computing Feynman diagrams in holomorphic (and mixed holomorphic-topological) field theories (see the next section for more on renormalization within the context of holomorphic QFT).
Such models are similar in spirit to Lie conformal algebras which use the notion of single-variable $\lambda$-bracket to encode the `singular part' of the data of a vertex algebra.

@@

@@

\owen{Somewhere we should mention Henriques' work connecting factorization algebras with conformal nets,
Daniel Bruegmann's thesis, and how the web of 2d CFT stuff should fit together.}

@@

For complex manifolds, and holomorphic factorization algebras on them, we consider an explicit example that generalizes a well-known concept in complex dimension one.
First, consider a \textit{real} one-dimensional \textit{topological} factorization algebra $\cA$.
If we identify $\cA$ with the associative algebra that it determines, then the factorization homology $\int_{S^1} \cA$ along the circle agrees with the Hochschild cohomology of the associative algebra $\cA$.
In other words, it is the home for all of the `traces' which the algebra $\cA$ admits.
It is well-known that the elliptic curve, or genus one, conformal blocks of a vertex algebra $\mathbb{V}$ is the natural `home' for the characters of the vertex algebra.
The most basic character of a conformal vertex algebra is the $q$-character defined in terms of a trace:
\beqn
\op{char}_{\mathbb{V}} (q) = \op{Tr}_{\mathbb{V}} (q^{L_0-c/24}) ,
\eeqn
where $L_0$ denotes the action of the vector field $z \del_z$.
The shift by the central charge guarantees that for $\mathbb{V}$ conformal that this expression has nice modularity properties.
From the point of view of factorization algebras, we can understand this trace in terms of Hochschild homology of the \textit{modes} algebra associated to $\mathbb{V}$ as introduced in \S\ref{??}.
Indeed, suppose we know that $\mathbb{V}$ corresponds to a holomorphic factorization algebra $\cF_{\mathbb{V}}$ on $\C$.
Then, if $r \colon \C^\times \to \RR_{>0}$ is the radial projection, we have discussed how to extract the modes algebra from the one-dimensional factorization algebra $r_* \cF_{\mathbb{F}}$.
If we further put this on the circle, we find Hochschild homology.
Thus, for every modulus $q$ one has
\beqn
\int_{E_q} \cF_{\mathbb{V}} = \int_{S^1} r_* \cF_{\mathbb{V}} = \op{Hoch}(r_* \cF_{\mathbb{V}}) .
\eeqn
For more on this perspective see \cite{GuiLi}.
From this point of view we hope it is clear how the factorization homology along an elliptic curve encodes characters in the sense of vertex algebras.

In higher dimensions we can build a similar picture.
The simplest so-called \textit{Hopf manifolds} are complex manifolds of the form
\beqn
H_{\mathbf{q}} = \left(\C^d - \{0\}\right) \slash \left((z_1,\ldots,z_d) \sim (q_1 z_1,\ldots,q_d z_d) \right)
\eeqn
where $\mathbf{q} = (q_1,\ldots,q_d)$ are complex numbers satisfying $0 < |q_i| < 1$.
For generic $\mathbf{q}$ such a complex manifold is diffeomorphic to $S^{2d-1} \times S^1$.
The factorization homology of a holomorphic factorization algebra $\cF$ along a Hopf surface of this kind can be identified with the Hochschild homology of the associative algebra of $S^{2n-1}$-modes:
\beqn
\int_{H_{\mathbf{q}}} \cF \simeq \op{Hoch}(r_* \cF)
\eeqn
where, again, $r\colon \C^d - \{0\} \to \RR_{>0}$ is the radius.
Thus, we see that the factorization homology $\int_{H_{\mathbf{q}}} \cF$ naturally encodes traces of the $S^{2n-1}$-mode algebra associated to $\cF$.
In \cite{SWchar}, we have shown how an important invariant called the \textit{supersymmetric index} (or supersymmetric partition function) of a supersymmetric theory along a sphere $S^{2n-1}$ be cast into this framework at the level the holomorphic twist.

@@

There are two diagrams that contribute to the gauge anomaly,
shown in Figure~\ref{fig:localization}:
\begin{itemize}
\item The pure gauge anomaly: it is proportional to the representation-theoretic factor $\op{Tr}_{ad}(X^3)$, $X \in \lie{g}$.
This expression vanishes when $\lie{g}$ is simple, e.g., for $\lie{g} = \lie{sl}(n)$).
\item The mixed anomaly: its edges (propagators) are labeled by the $\beta\gamma$ fields,
and it is proportional to the representation-theoretic factor $\op{Tr}_V(X^3)$, $X \in \lie{g}$.
\end{itemize}
These each come from ways of labeling the vertices and edges of the wheel (or polygon) diagram with three vertices.
We have described the representation-theoretic factor but there is also an analytic factor.
\owen{Should we just write it?}

@@

\owen{There isn't a $K$-edge on the wheel?}

@@

		\begin{scope}[shift={(6,0)}]
				\coordinate (O) at (0,0);
		\node[draw] (ci) at (O) [circle through=(right:1)] {};
		%\draw[fill=black] (0,0) circle (1cm);
		%\draw[fill=white] (0,0) circle (0.99cm);
		%\draw[line width=0.35mm,red] ++(145:0.995) arc (145:215:0.995);
		%\node[red] at (180:0.7) {$K_\epsilon$};
		\draw[fermion](45:2) -- (45:1);
		\draw[fermion](180:2) -- (180:1);
		\draw[fermion](315:2) -- (315:1);
		\node at (45:2.2) {$A$};
		\node at (180:2.2) {$A$};
		\node at (315:2.2) {$A$};
		\node at (130:1.4) {$P_{\beta\gamma}$};
		\node at (270:1.4) {$P_{\beta\gamma}$};
		\node at (0:1.4) {$P_{\beta\gamma}$};
	    	\clip (0,0) circle (1cm);
 \end{scope}
\end{tikzpicture}
\caption{Diagrams which contribute to the anomaly in holomorphic gauge theory on $\C^2$}
\label{fig:localization}
\end{center}
\end{figure}

@@

We begin this section by describing the holomorphic twist of $\cN=1$ supersymmetric Yang--Mills theory,
and then we offer a holomorphic version of Seiberg duality, 
initially conjectured by Richard Eager.
We also explain how this duality, when compactified along Riemann surfaces,
should produce equivalences between different kinds of 2-dimensional B-model theories;
these reductions are related to work of Hori and Hori--Tong \owen{citations} and its mathematical development by Ed Segal, \owen{list others}.

>>>>

\begin{rmk}
In the $d=1$ case there is a lovely exploration of how the highly developed theory of vertex algebras relates to this perspective in the pair of papers by Bruegmann \cite{}.
For a discussion of how conformal nets --- an alternative formalism for chiral CFT --- relates to our perspective, see \cite{Henriques}.
\end{rmk}

Before jumping into the main story, we mention a motivation for talking about $E_d$- and $\PDisc_d$-algebras in parallel.
By dealing with both, we are suggesting, of course, that there is a natural analogy and hence a useful transfer of intuition between the settings of topological and holomorphic field theories.
But there is a stronger, technical relationship that we'll develop via a version of dimensional reduction.
From each holomorphic field theory in complex dimension $d$, 
we'll extract an $E_d$-algebra.
This construction generalizes the relationship between the Kac-Moody vertex algebra of a Lie algebra $\fg$ at level $k$ and the central extension of the loop algebra~$L\fg$.
We expect this correspondence to bear rich fruit.

There is an explicit model for an algebra over the cohomology of the $E_d$-operad called a $P_d$-algebra.
Essentially, these are graded commutative algebras with a Poisson bracket of cohomological degree $-d+1$.
In \cite{CostelloScheimbauer} a parallel model for the cohomology of the holomorphic $\PDisc_d$ colored operad was proposed, 
though axioms were not written down.
In recent work, similar models to that one have appeared,
using the nomenclature of $\lambda$-brackets \cite{BGKWY,GKW},
following Kac's approach in the $d=1$ setting. 
The multivariable $\lambda$-brackets are similar in spirit to Lie conformal algebras,
which use the notion of single-variable $\lambda$-bracket to encode the ``singular part'' of the data of a vertex algebra.
These papers use this algebraic structure to give a beautiful proposal for computing Feynman diagrams in holomorphic (and mixed holomorphic-topological) field theories (see Section~\ref{sec: renorm} for more on renormalization within the context of holomorphic QFT).

@@

A recent survey of factorization algebras, with a focus on their role in mathematical physics, can be found in~\cite{CosGwEMP}.

@@

\owen{Not sure what to put here, honestly. Maybe it should be removed.}

@@

For complex manifolds, and holomorphic factorization algebras on them, we consider an explicit example that generalizes a well-known concept in complex dimension one.
First, consider a \textit{real} one-dimensional \textit{topological} factorization algebra $\cA$.
If we identify $\cA$ with the associative algebra that it determines, then the factorization homology $\int_{S^1} \cA$ along the circle agrees with the Hochschild cohomology of the associative algebra $\cA$.
In other words, it is the home for all of the `traces' which the algebra $\cA$ admits.
It is well-known that the elliptic curve, or genus one, conformal blocks of a vertex algebra $\mathbb{V}$ is the natural `home' for the characters of the vertex algebra.
The most basic character of a conformal vertex algebra is the $q$-character defined in terms of a trace:
\beqn
\op{char}_{\mathbb{V}} (q) = \op{Tr}_{\mathbb{V}} (q^{L_0-c/24}) ,
\eeqn
where $L_0$ denotes the action of the vector field $z \del_z$.
The shift by the central charge guarantees that for $\mathbb{V}$ conformal that this expression has nice modularity properties.
From the point of view of factorization algebras, we can understand this trace in terms of Hochschild homology of the \textit{modes} algebra associated to $\mathbb{V}$ as introduced in \S\ref{??}.
Indeed, suppose we know that $\mathbb{V}$ corresponds to a holomorphic factorization algebra $\cF_{\mathbb{V}}$ on $\C$.
Then, if $r \colon \C^\times \to \RR_{>0}$ is the radial projection, we have discussed how to extract the modes algebra from the one-dimensional factorization algebra $r_* \cF_{\mathbb{F}}$.
If we further put this on the circle, we find Hochschild homology.
Thus, for every modulus $q$ one has
\beqn
\int_{E_q} \cF_{\mathbb{V}} = \int_{S^1} r_* \cF_{\mathbb{V}} = \op{Hoch}(r_* \cF_{\mathbb{V}}) .
\eeqn
For more on this perspective see \cite{GuiLi}.
From this point of view, the factorization homology along an elliptic curve encodes characters in the sense of vertex algebras.

In higher dimensions we can build a similar picture.
The simplest so-called \textit{Hopf manifolds} are complex manifolds of the form
\beqn
H_{\mathbf{q}} = \left(\C^d - \{0\}\right) \slash \left((z_1,\ldots,z_d) \sim (q_1 z_1,\ldots,q_d z_d) \right)
\eeqn
where $\mathbf{q} = (q_1,\ldots,q_d)$ are complex numbers satisfying $0 < |q_i| < 1$.
For generic $\mathbf{q}$ such a complex manifold is diffeomorphic to $S^{2d-1} \times S^1$.
The factorization homology of a holomorphic factorization algebra $\cF$ along a Hopf surface of this kind can be identified with the Hochschild homology of the associative algebra of $S^{2n-1}$-modes:
\beqn
\int_{H_{\mathbf{q}}} \cF \simeq \op{Hoch}(r_* \cF)
\eeqn
where, again, $r\colon \C^d - \{0\} \to \RR_{>0}$ is the radius.
Thus, we see that the factorization homology $\int_{H_{\mathbf{q}}} \cF$ naturally encodes traces of the $S^{2n-1}$-mode algebra associated to $\cF$.
In \cite{SWchar}, it is shown how an important invariant called the \textit{supersymmetric index} (or supersymmetric partition function) of a supersymmetric theory along a sphere $S^{2n-1}$ be cast into this framework at the level the holomorphic twist.

@@

There are two diagrams that contribute to the gauge anomaly,
shown in Figure~\ref{fig:localization}.
\owen{There isn't a $K$-edge on the wheel?}
These define functions on the space of fields:
\begin{itemize}
\item The pure gauge anomaly: it is proportional to the representation-theoretic factor $\op{Tr}_{ad}(X^3)$ for  $X \in \lie{g}$, which is an invariant function of~$\fg$.
This expression vanishes when $\lie{g}$ is simple, e.g., for $\lie{g} = \lie{sl}(n)$.
\item The mixed anomaly: its edges (propagators) are labeled by the $\beta\gamma$ fields,
and it is proportional to the representation-theoretic factor $\op{Tr}_V(X^3)$ for $X \in \lie{g}$.
\end{itemize}
These each come from ways of labeling the vertices and edges of the wheel (or polygon) diagram with three vertices.
We have described the representation-theoretic factor but there is also an analytic factor.
\owen{Should we just write out the formulas? What we write is kind of hard to parse, as it is shorthand. I can imagine a person asking "proportional ... how?"}

@@

@@

		\begin{scope}[shift={(6,0)}]
				\coordinate (O) at (0,0);
		\node[draw] (ci) at (O) [circle through=(right:1)] {};
		%\draw[fill=black] (0,0) circle (1cm);
		%\draw[fill=white] (0,0) circle (0.99cm);
		%\draw[line width=0.35mm,red] ++(145:0.995) arc (145:215:0.995);
		%\node[red] at (180:0.7) {$K_\epsilon$};
		\draw[fermion](45:2) -- (45:1);
		\draw[fermion](180:2) -- (180:1);
		\draw[fermion](315:2) -- (315:1);
		\node at (45:2.2) {$A$};
		\node at (180:2.2) {$A$};
		\node at (315:2.2) {$A$};
		\node at (130:1.4) {$P_{\beta\gamma}$};
		\node at (270:1.4) {$P_{\beta\gamma}$};
		\node at (0:1.4) {$P_{\beta\gamma}$};
	    	\clip (0,0) circle (1cm);
 \end{scope}
\end{tikzpicture}
\caption{Diagrams that contribute to the anomaly in holomorphic gauge theory on~$\C^2$}
\label{fig:localization}
\end{center}
\end{figure}

@@

We begin this section by describing the holomorphic twist of $\cN=1$ supersymmetric Yang--Mills theory,
and then we offer a holomorphic version of Seiberg duality, 
initially conjectured by Richard Eager.
We end by sketching how this duality, when compactified along Riemann surfaces,
should produce equivalences between different kinds of 2-dimensional B-model theories;
these reductions are related to work of Hori and Hori--Tong \owen{citations} and its mathematical development by Ed Segal, \owen{list others}.
